### 2021-05-19 project_ashish_monograph/LHC.tex
monograph updated

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity \cite{cmsrun1lumi}. During Run 2, LHC instantaneous luminosity reached a peak value of $2.1 \times 10^{34}  \: cm^{-2} s^{-1}$ with $163 \: fb^{-1}$ integrated luminosity as shown in Fig. 2 \cite{CMS:2018elu}. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase 2 HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value \cite{Dainese:2019rgk}. \\

>>>>

Luminosity is a key performance parameter of an accelerator as it is a measure of number of collisions. Luminosity can be categorized into two types: instantaneous and integrated luminosity. Instantaneous luminosity is defined as the number of collisions occuring per second in per centimeter square area while integrated luminosity is obtained by integrating instantaneous luminosity over a time interval. Both are very important variables as they are used to obtain the number of collision events. During Run 1, peak value of LHC instantaneous luminosity was $0.77 \times 10^{34} \: cm^{-2} s^{-1} $ with $30 \: fb^{-1}$ integrated luminosity \cite{cmsrun1lumi}. During Run 2, LHC instantaneous luminosity reached a peak value of $2.1 \times 10^{34}  \: cm^{-2} s^{-1}$ with $163 \: fb^{-1}$ integrated luminosity as shown in Fig. 2 \cite{CMS:2018elu}. At the end of Run 3, we expect to achieve an integrated luminosity  of $300 \:fb^{-1}$ while Phase II HL-LHC is expected to collect $3000 \:fb^{-1}$ of data as shown in Fig. 3 thereby increasing luminosity by a factor of 10 beyond the LHC's design value \cite{Dainese:2019rgk}. \\

### 2021-05-19 project_ashish_monograph/PCC.tex
monograph updated

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}.

>>>>

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}. \\

1. Run2018A \\

Run number ranges from 315255 to 316995 \\
Number of runs per period is 146 \\

2. Run2018B\\

Run number ranges from 317080 to 319311\\
Number of runs per period is 132\\

3. Run2018C\\

Run number ranges from 319337 to 320065\\
Number of runs per period is 88 \\

The veto lists corresponds to second Run2018D and select each detector part L2, L3, L4, D1, D2, D3 \cite{vetolist}. \\

1. vetolateRunDlowcuttight (Number of vetoed modules 1387) \\
2. vetolateRunDlowcuttightB1 (Number of vetoed modules 1850)\\
3. vetolateRunDlowcuttightB2 (Number of vetoed modules 1766)\\
4. vetolateRunDlowcuttightB3 (Number of vetoed modules 1641)\\
5. vetolateRunDlowcuttightF1 (Number of vetoed modules 1813)\\
6. vetolateRunDlowcuttightF2 (Number of vetoed modules 1799)\\
7. vetolateRunDlowcuttightF3 (Number of vetoed modules 1800)\\

### 2021-05-19 project_ashish_monograph/results.tex
monograph updated

The stability of luminosity determination using pixel detector based on PCC method is investigated using late Run2018D veto list.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./316993.png}
  \caption{Top: Bunch crossing id (bcid) and PCC lumi as a function of lumi section (1 lumi section is 23.36s). Bottom: PCC lumi as a function of lumi section (late RunD veto list is used) for run 2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ls_lumi.png}
  \caption{PCC luminosity ($\mu b^{-1}$) as a function of lumi section (1 lumi section is 23.36s).}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./319270.png}
  \caption{Top: Bunch crossing id (bcid) and PCC lumi as a function of lumi section. Bottom: PCC lumi as a function of lumi section (late RunD veto list is used) for run 2018B.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./320061.png}
  \caption{Top: Bunch crossing id (bcid) and PCC lumi as a function of lumi section. Bottom: PCC lumi as a function of lumi section (late RunD veto list is used) for run 2018C.}
  \label{fig:CMS}
\end{figure}

>>>>

The stability of CMS integrated luminosity for Phase I pixel detector calculated using PCC method is investigated using late Run2018D veto list.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./316993.png}
  \caption{Top: Bunch crossing id (bcid) and PCC luminosity as a function of lumi section (1 lumi section is 23.36s). Bottom: PCC lumi as a function of lumi section using late run 2018D veto list for run 2018A.}
  \label{fig:CMS}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./ls_lumi.png}
  \caption{PCC luminosity ($\mu b^{-1}$) as a function of luminosity section (1 lumi section is 23.36s).}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.52\columnwidth]{./319270.png}
  \caption{Top: Bunch crossing id (bcid) and PCC luminosity as a function of lumi section. Bottom: PCC lumi as a function of lumi section using late run 2018D veto list for run 2018B.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./320061.png}
  \caption{Top: Bunch crossing id (bcid) and PCC luminosity as a function of lumi section. Bottom: PCC lumi as a function of lumi section using late run 2018D veto list for run 2018C.}
  \label{fig:CMS}
\end{figure}

### 2021-05-19 project_ashish_monograph/syst.tex
monograph updated

Systematic uncertainties in the bunch overlap widths along x and y directions measured during vdM scans, $\sigma_{vis}$ and CMS luminosity can arise from length scale calibration, orbit drift, x-y correlations of bunch proton density function, emittance growth, bunch-bunch deflection, dynamic $\beta$ which tells us whether bunches are narrow or wide, bunch current calibration, ghosts and satellites, scan to scan variation and bunch to bunch variation. Systematic uncertainty during physics runs (high pileup) can be due to cross-detector consistency, afterglow effects, linearity and CMS deadtime.  \\

1. Length scale calibration:  Value of the separation derived from currents in the corrector magnets, calibrated against central tracking system and compare luminous region information (vertices distribution) versus nominal separation. \\

2. Emittance growth: Bunches size are known to increase in size during the fill. Effect is sizable during the ~30 minutes, Bias almost negligible if emittance grows linearly with time and if measurement are made in between X and Y scans. \\

3. Orbit drift: The beams can slowly drift in the transverse plane. This effect is typically small but can be harmful in some cases. Drift of the orbit estimated by BPM measurements taken between scans and extrapolated to interaction point. \\

4. xy-correlations: The vdM scan method assumes that the bunch proton density function is factorizable into independent x- and y-dependent terms which is not fully true and can lead to a biased estimate of the bunch overlap area. In order to measure this effect, special beam imaging scans are conducted. In these scans, the measured vertex position distributions are used to derive the bunch proton densities, which can then be used to estimate a correction to the visible cross section $\sigma_{vis}$. \\

5. Linearity: \\

6. Stability:  \\

7. Bunch-Bunch Effects: \\

Dipolar kick (bunch-bunch deflection): Repulsive force deflects beams affecting nominal separation depends on the separation itself, the beam width and the current. \\

 Quadrupolar (de)focusing (dynamic $\beta$): Effective beam width is modified depending on the separation which in turn affects the peak luminosity (rate) value. \cite{CMS-PAS-LUM-18-002} \\

>>>>

Systematic uncertainties in the beam overlap widths along x and y directions measured during vdM scans, $\sigma_{vis}$ and CMS luminosity can arise from length scale calibration, orbit drift, XY correlations of beam proton density distribution function, emittance growth, beam-beam deflection, dynamic $\beta$ which tells us whether bunches are narrow or wide, beam current calibration, ghosts and satellites, scan to scan variation and bunch to bunch variation. Systematic uncertainty during physics runs (high pileup) can be due to cross-detector consistency, afterglow effects, linearity and CMS deadtime. \\

\textbf{Systematics uncertainties during vdM scans or in $\bf{\sigma_{vis}}$:} \\

1. Length scale calibration: Acurracy of the beam positions from the LHC magnetic currents is limited that is why length scale calibration is done in which beams are moved in forward and backward directions at constant separation. A linear fit between the measured position and nominal position is plotted and deviation of the slope of this curve from one is used as correction. \\

2. Emittance growth: Beam size is known to increase during the fill which affects the cluster rate profile amplitude, its width and luminosity. Their variation is studied as a function of time and a linear fit is performed to obtain correction. Effect is sizable during a period of 30 minutes. \\

3. Orbit drift: Beams can slowly drift from their nominal positions in the transverse plane which causes beam separation to vary over a period of vdM scan step. LHC beam position monitoring systems are used to correct for beam separation. \\

4. XY correlations: The vdM scan method assumes that the shape of the bunch proton density function is factorizable into independent X and Y dependent terms which is not fully true and can lead to a biased estimate of the bunch overlap area $\Sigma_x \Sigma_y$. In order to measure this effect, special beam imaging scans are conducted in which one beam is fixed and other beam is moved across. In these scans, the measured vertex position distributions are used to derive the beam proton densities where Super Double Gaussian function is used as best fit and correction to the visible cross section $\sigma_{vis}$ is estimated. \\

5. Beam-Beam Effects: \\

Dipolar kick (beam-beam deflection): Repulsive force due to beam's electric field deflects other beam which affect nominal separation, the beam width and beam current. \\
Quadrupolar defocusing (dynamic $\beta$): Beam can get defocused by other beam's magnetic quadrupole field. Effective beam width is modified depending on the separation which in turn affects the peak instantaneous luminosity value because collision rate is decreased by 10 \%. \\

\newpage
\textbf{Systematics uncertaintities during physics runs (high pileup)}:\\

6. Non-Linearity: extrapolation of $\sigma_{vis}$ from vdM scans to high pileup physics data-taking periods can cause it to vary with per bunch instantaneous luminosity and pileup if the relation between luminosity and pileup is not linear.                               \\

7. Stability: Instability can arise from variation of measured cluster rate over time for different subdetectors in pixel detector and change of detector conditions over time for example detector noise, aging effects, radiation damage, etc. Results for PCC luminosity stability for 2018 data is described in the next section.                               \\

8. Cross-detector comparisons: After corrections of out-of-time effects and efficiency changes monitored by emittance scans, residual effects are evaluated by cross-detector comparison.                    \\

8. Afterglow effects: A small out of time response appears in unfilled bunches after colliding bunches. These effects are described in detail in the previous section \cite{CMS-PAS-LUM-18-002}.

### 2021-05-19 project_ashish_monograph/upgrade.tex
monograph updated

Instantaneous luminosity determination using PCC method for Phase II HL-LHC can be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth. TEPX Disk 4 Ring 1 layout is shown in Fig. 29. Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 31 and Fig. 32 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 31. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences.  \cite{Collaboration:2706512} \\

@@

\newpage Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. Linearity results for simulated TEPX clusters and coincidences from low to high pileup values are shown from Fig. 33 to Fig. 50. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./fourlayers.png}
  \caption{ \onehalfspacing Fours layers of one double disk of TEPX showing module arrangement in rings for all layers.}
  \label{fig:CMS}
\end{figure}

>>>>

Instantaneous luminosity determination using PCC method for Phase II HL-LHC will be based on counting the number of clusters or coincidences. The innermost ring of the last disk of TEPX (D4R1) is located at 2.65 m away from the interaction point that is beyond the tracking acceptance ($|\eta| = 4$) and as this region has few tracking points, it can be solely used for the purpose of luminosity measurement by using the full available trigger rate and bandwidth. TEPX Disk 4 Ring 1 layout is shown in Fig. 29. Two fold coincidences are those hits which are created by the module overlap regions between various layers of one TEPX double disk. Two fold coincidences are better way to distinguish between a real hit and random electrical noise. They are more likely to be real hit than random electrical noise. Two fold coincidence in $\phi$ will involve modules overlapping in the same ring in front and back layers of one double disk as shown in top part of Fig. 31 and Fig. 32 while two fold coincidence in r will require modules overlapping between successive rings in the front (L1 $\&$ L2) and back layers (L3 $\&$ L4) of one double disk as shown in bottom part of Fig. 31. Luminosity calculated based on counting coincidences has an advantage over clusters that afterglow effects are tiny in the case of coincidences.  \cite{Collaboration:2706512} \cite{brilsim} \cite{brilsim1}\cite{brilsim2}\\

@@

Linearity indicates that the PCC visible cross section $\sigma_{vis}$ does not depend on the per bunch instantaneous luminosity (ideal luminometer). In ideal scenario, $\sigma_{vis}$ is not dependent on pileup, but this a potential problem with the luminometer. Linearity results for simulated TEPX clusters and coincidences from low to high pileup values are shown from Fig. 33 to Fig. 50. TEPX luminometer proposed for Phase II shows excellent linearity over entire pileup range. A non-linear relation $<N_{cluster}> = \alpha (PU)^{\gamma}, \gamma \neq 1$ would add non-linear terms in the rate equation $R = \sigma_{vis} L_{inst}$  and cause $\sigma_{vis}$ to vary with the per bunch instantaneous luminosity and pileup values.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=1 \columnwidth]{./fourlayers.png}
  \caption{ \onehalfspacing Fours layers of one double disk of TEPX showing module arrangement in rings for each layer.}
  \label{fig:CMS}
\end{figure}
